Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-07-23 concepts.tex
AA: non ho fatto niente di che. Riprendo stasera

Let $S$ be a parameterized system with $n$ parameters. The generic
parameter $p_i, \ i \in \{1,2,\ldots,n\}$ can take any value in
the set $V_i$. A {\em configuration} $\mathbf{c}$ of the system
$S$ is a $n$-tuple $\langle v_1,v_2,\ldots,v_n \rangle$ in which
$v_i \in V_i$ is the value fixed for the parameter $p_i$. The {\em
configuration space} (or {\em design space}) of $S$ [which we will
indicate as $\mathcal{C}(S)$] is the complete range of possible
configurations [$\mathcal{C}(S) = V_1 \times V_2 \times \ldots
\times V_n$]. Naturally not all the configurations of
$\mathcal{C}(S)$ can really be mapped on $S$. We will call the set
of configurations that can be physically mapped on $S$ the {\em
feasible configuration space} of $S$ [and indicate it as
$\mathcal{C}^*(S)$].

@@

Given a system $S$, an application $b \in \mathcal{B}$ and two
configurations $\mathbf{c}', \mathbf{c}'' \in \mathcal{C}^*(S)$,
$\mathbf{c}'$ is said to {\em dominate} (or {\em eclipse})
$\mathbf{c}''$, and is indicated as $\mathbf{c}' \succ
\mathbf{c}''$, if given $\mathbf{o}'=E(\mathbf{c}', b)$ and
$\mathbf{o}''=E(\mathbf{c}'', b)$ it results that $\mathbf{o}'
\leq \mathbf{o}''$ and $\mathbf{o}' \neq \mathbf{o}''$. Where
vector comparisons are interpreted component-wise and are true
only if all of the individual comparisons are true ($o'_i \leq
o''_i \ \forall \ i = 1,2,\ldots,m$).

\begin{definition}[Pareto Set]
\label{pers02.def:Pareto-set}
The {\em Pareto-optimal set} of $S$ for the application $b$ is the
set:
\[ \mathcal{P}(S,b) = \{ \mathbf{c} \in \mathcal{C}^*(S) \ : \nexists \ \mathbf{c}' \in \mathcal{C}^*(S), \mathbf{c}' \succ \mathbf{c} \} \]
\end{definition}
that is, the set of configurations $\mathbf{c} \in
\mathcal{C}^*(S)$ not dominated by any other configuration.
Pareto-optimal configurations are configurations belonging to the
Pareto-optimal set and the {\em Pareto-optimal front} is the image
of the Pareto-optimal configurations, i.e. the set:
\[ \mathcal{P}_{F}(S,b) = \{ \mathbf{o} : \mathbf{o} = E(\mathbf{c},b), \mathbf{c} \in \mathcal{P}(S,b) \} \]

@@

On the contrary, we call ``parameter space representation'' of the
Pareto set the representation in which each configuration $\mathbf{p}$
is simply represented as a point of $P_{1}\times\dots\times P_{n}$
space. We will investigate the relation between the ``objective space
representation'' and the ``parameter space representation''. We
are particularly interested in the relation between the evolution
of the objective space representation of the Pareto set and the evolution
of its parameter space representation.

@@

In the design process, there are often a lot of ``local decisions''
in which the designer can take advantage of well-known results or
heuristics. For example, in a VLIW architecture consider a configuration
with very few registers. It is well-known that, in such a condition,
increasing the number of ALUs doesn't lead to tangible better performances
but leads to an increase in area occupation. An intelligent exploration
algorithm wouldn't waste time evaluating configurations with many
ALUs and few registers. If $p_{1}$ is the parameter ``number of
registers'' and $p_{2}$ is the parameter ``number of ALUs'', we
can say that the region:
\[
R=\left\{ \left.\left(v_{1},v_{2}\right)\in P_{1}\times P_{2}\right|v_{1}<s_{1},v_{2}>s_{2}\right\} 
\]
 (where $s_{1}$ and $s_{2}$ are some threshold values) is of little
interest. This example shows that some ``dependency relation'' between
parameters may exist and can be exploited to guide the exploration:
the dependencies between parameters can suggest to the designer that,
if a parameter lies in a range, there are ranges of the other there
aren't worth exploring. Intuitively, the Cartesian product of these
ranges is an ``uninteresting region''.

>>>>

Let $S$ be a parameterized system with $n$ parameters. The generic
parameter $p_i, \ i \in \{1,2,\ldots,n\}$ can take any value in
the set $V_i$. A {\em configuration} $\mathbf{c}$ of the system
$S$ is a $n$-tuple $\langle v_1,v_2,\ldots,v_n \rangle$ in which
$v_i \in V_i$ is the value fixed for the parameter $p_i$. The {\em
configuration space} (or {\em design space}) of $S$ [which we will
indicate as $\mathcal{C}(S)$] is the complete range of possible
configurations [$\mathcal{C}(S) = V_1 \times V_2 \times \ldots
\times V_n$]. Naturally, not all the configurations of
$\mathcal{C}(S)$ can really be mapped on $S$, but only a subset 
is feasible. We call this subset {\em
feasible configuration space} $S$ and indicate it as
$\mathcal{C}^*(S)$].

@@

Given a system $S$, an application $b \in \mathcal{B}$ and two
configurations $\mathbf{c}', \mathbf{c}'' \in \mathcal{C}^*(S)$,
$\mathbf{c}'$ is said to {\em dominate} (or {\em eclipse})
$\mathbf{c}''$, if, given $\mathbf{o}'=E(\mathbf{c}', b)$ and
$\mathbf{o}''=E(\mathbf{c}'', b)$, it results that $\mathbf{o}'
\leq \mathbf{o}''$ and $\mathbf{o}' \neq \mathbf{o}''$, where
vector comparisons are interpreted component-wise and are true
only if all of the individual comparisons are true ($o'_i \leq
o''_i \ \forall \ i = 1,2,\ldots,m$). To indicate that 
$\mathbf{c}'$ dominates $\mathbf{c}''$, we use the notation
$\mathbf{c}' \succ \mathbf{c}''$.

\begin{definition}[Pareto Set]
\label{pers02.def:Pareto-set}
The {\em Pareto-optimal set} of $S$ for the application $b$ is the
set:
\[ \mathcal{P}(S,b) = \{ \mathbf{c} \in \mathcal{C}^*(S) \ | \nexists \ \mathbf{c}' \in \mathcal{C}^*(S), \mathbf{c}' \succ \mathbf{c} \} \]
\end{definition}
that is, the set of configurations $\mathbf{c} \in
\mathcal{C}^*(S)$ not dominated by any other \comment{feasible}
 configuration.
The configurations belonging to the Pareto-optimal set are called \emph{Pareto-optimal configurations}, while the {\em Pareto-optimal front} is their image, i.e. the set:
\[ \mathcal{P}_{F}(S,b) = \{ \mathbf{o} | \mathbf{o} = E(\mathbf{c},b), \mathbf{c} \in \mathcal{P}(S,b) \} \]

@@

\comment{Mi sembra solo un cambio di notazione. }
On the contrary, we call ``parameter space representation'' of the
Pareto set the representation in which each configuration $\mathbf{p}$
is simply represented as a point of $P_{1}\times\dots\times P_{n}$
space. We will investigate the relation between the ``objective space
representation'' and the ``parameter space representation''. We
are particularly interested in the relation between the evolution
of the objective space representation of the Pareto set and the evolution
of its parameter space representation.

@@

In the design process, there are often a lot of ``local decisions''
in which the designer can take advantage of well-known results or
heuristics. For example, if we consider a VLIW architecture in a configuration with very few registers, it is well-known that increasing the number of ALUs does not bring tangible better performances
but only cause an increase in area occupation. An intelligent exploration
algorithm would not waste time evaluating configurations with many
ALUs and few registers. If $p_{1}$ is the parameter ``number of
registers'' and $p_{2}$ is the parameter ``number of ALUs'', we
can say that the region:
\[
R=\left\{ \left.\left(v_{1},v_{2}\right)\in P_{1}\times P_{2}\right|v_{1}<s_{1},v_{2}>s_{2}\right\} 
\]
 (where $s_{1}$ and $s_{2}$ are some threshold values) is of modest
interest. This example shows that some ``dependency relation'' between
parameters may exist and can be exploited to guide the exploration:
the dependencies between parameters can suggest to the designer that,
if a parameter lies in a range, there are ranges of the other there
aren't worth exploring. Intuitively, the Cartesian product of these
ranges is an ``uninteresting region''.

### 2014-07-23 introduction.tex
AA: non ho fatto niente di che. Riprendo stasera

\comment{Potrebbe essere pompato di piu' con qualche cosa piu' d'impatto? }

\section{Related work and Contribution}
\label{sec:Related-work}
\comment{Preferisci che i related work siano in un'unica sezione con Intro?}
Many different design space exploration algorithms have been proposed
in literature.  The motivations of an exploration
algorithm are rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

@@

The paper is organized as follows: in
Section~\ref{sec:statement_of_the_problem} we formally introduce the
main concepts and the definitions required to apply the proposed strategy. Next in
Section~\ref{sec:algorithm} we present the PS algorithm in detail.
Finally in Section~\ref{sec:ee} we show a case study involving the
exploration of the hardware/software parameters of a VLIW
architecture, performing a qualitative and quantitative comparison of
PS agains the state-of-art of multiobjective genetic based approaches.

>>>>

\section{Related work and Contribution}
\label{sec:Related-work}
Many different design space exploration algorithms have been proposed
in literature.  The motivations of an exploration
algorithm are rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

@@

The paper is organized as follows: in
Section~\ref{sec:statement_of_the_problem} we formally introduce the
main concepts and the definitions required to apply the proposed strategy. Next in
Section~\ref{sec:algorithm} we present the PS algorithm in detail.
Finally in Section~\ref{sec:ee} we show a case study involving the
exploration of the hardware/software parameters of a Very Long
 Instruction Word (VLIW) architecture, performing a qualitative
 and quantitative comparison of
PS agains the state-of-art of multiobjective genetic based approaches.

### 2014-07-23 paper.tex
AA: non ho fatto niente di che. Riprendo stasera

\documentclass[final,journal,letterpaper]{IEEEtran}

@@

\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{amsthm}

@@

\usepackage{listings}
\usepackage{listings}

@@

%\renewcommand{\baselinestretch}{0.98}\normalsize
%------------------------------------------------------------------------------
\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}
\theoremstyle{definition}
\newtheorem{defn}{\protect\definitionname}
\theoremstyle{remark}
\newtheorem{rem}{\protect\remarkname}

%\newtheorem{defn}[thm]{\protect\definitionname}
%\theoremstyle{remark}
%\newtheorem{rem}[thm]{\protect\remarkname}
%\makeatother

@@

\begin{document}

\author{Vincenzo~Catania, Andrea Araldo~and~Davide~Patti% 
\thanks{The authors are with the Dipartimento di Ingegneria
Elettrica, Elettronica ed Informatica, Universit\`a di Catania, Italy
(email: \{vincenzo.catania,aaraldo,davide.patti\}@dieei.unict.it).}}

\title{Parameter Space Representation of Pareto Front to
Explore Hardware-Software Dependencies}
%ATTENZIONE: in questo testo io mi sono riferito a ``Pareto set''
%come un sottoinsieme dello spazio degli obiettivi (vedi, ad es., descrizione
%dell'algoritmo). E' più corretto forse, invece, intendere il Pareto
%set come un sottoinsieme dello spazio delle configurazioni.

\maketitle
\begin{abstract}
Embedded Systems Design often requires conflicting objectives to be
optimized with an appropriate choice of hardware/software parameters
. Interdependencies among these parameters can make this task even
more difficult, since an a-priori knowledge of them would  be required
to develop an analytical model of the system. In this work we present
PS, a multiobjective exploration strategy that introduces the notion
of Parameters Space Representation. After a detailed formal
description of the algorithm and its concepts, we show a
hardware/software exploration case study. Qualitative and quantitative
comparisons of PS against a well-know multiobjective genetic approach
demonstrate that, while not outperforming it in terms of dominance, the
proposed approach can trade the uniformity and granularity
qualities of the solutions found for obtaining larger Pareto sets,
thus representing a further choice for designer when objective constrains allow it.
\end{abstract}

%------------------------------------------------------------------------------
%\IEEEkeywords{...}

\begin{IEEEkeywords}
DSE, VLIW, MultiObjective
\end{IEEEkeywords}
%------------------------------------------------------------------------------
\input{introduction.tex}
\input{concepts.tex}
\input{algorithm.tex}
\input{experiments.tex}
%------------------------------------------------------------------------------

\section{Conclusions}
This is the end

@@

\bibliographystyle{IEEEtran} 
\bibliography{bibliography}

>>>>

%\documentclass[final,journal,letterpaper]{IEEEtran}
\documentclass[prodmode,acmtecs]{acmsmall}

@@

\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{textcomp}
\usepackage{mathrsfs}
%\usepackage{amsthm}

@@

\usepackage{listings}

@@

%\renewcommand{\baselinestretch}{0.98}\normalsize
%------------------------------------------------------------------------------
%\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
%\theoremstyle{plain}
%\newtheorem{thm}{\protect\theoremname}
%\theoremstyle{definition}
%\newtheorem{defn}{\protect\definitionname}
%\theoremstyle{remark}
%\newtheorem{rem}{\protect\remarkname}

@@

\begin{document}
% Page heads
\markboth{V. Catania et al.}{Parameter Space Representation of Pareto Front to
Explore Hardware-Software Dependencies}

\title{Parameter Space Representation of Pareto Front to
Explore Hardware-Software Dependencies}
\author{VINCENZO CATANIA
\affil{University of Catania}
ANDREA ARALDO
\affil{University of Catania}
DAVIDE PATTI
\affil{University of Catania}
}

%\maketitle
\begin{abstract}
Embedded Systems Design often requires conflicting objectives to be
optimized with an appropriate choice of hardware/software parameters .
Interdependencies among these parameters can make this task even more
difficult, since an a-priori knowledge of them would  be required to
develop an analytical model of the system. In this work we present PS,
a multiobjective exploration strategy that introduces the notion of
Parameters Space Representation. After a detailed formal description
of the algorithm and its concepts, we show a hardware/software
exploration case study. Qualitative and quantitative comparisons of PS
against a well-know multiobjective genetic approach demonstrate that,
while not outperforming it in terms of dominance, the proposed
approach can trade the uniformity and granularity qualities of the
solutions found for obtaining larger Pareto sets, thus representing a
further choice for designer when objective constrains allow it.
\end{abstract}

\category{D.2.2}{Design Tools and Techniques}{}

\terms{Design, Algorithms, Performance}

\keywords{DSE, MultiObjective Optimization, Genetic Algorithms}

\acmformat{Vincenzo Catania, Andrea Araldo and Davide Patti, 2014. Parameter Space Representation of Pareto Front to
Explore Hardware-Software Dependencies.}

\begin{bottomstuff}
Author's addresses: V. Catania, DIEEI University of Catania;
A. Araldo, DIEEI University of Catania;
D. Patti, DIEEI University of Catania.
\end{bottomstuff}

\maketitle
%------------------------------------------------------------------------------
\input{introduction.tex}
\input{concepts.tex}
\input{algorithm.tex}
\input{experiments.tex}
%------------------------------------------------------------------------------

\section{Conclusions}
In this work we presented a multiobjective strategy which 
introduces the concept of Parameter Space representation of Pareto
Set. A case study of a VLIW architecture involving strong
hardware/software dependencies has been carried out to evaluate
effectiveness with different amount of simulations budget.
A qualitative/quantitative comparison against a widespread multiobjective genetic
approach showed how the proposed strategy can result in a different
distribution of Pareto points which trades solutions granularity with
improved pareto extension.

@@

\bibliographystyle{ACM-Reference-Format-Journals} 
\bibliography{bibliography}
